Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeast Taiwan

### 2019-10-04 analysis/paper/paper.Rmd
add a paragraph about Kiwulan in the Materials section

@@

# Research area
Kiwulan is located at the north part of Yilan plain in northeast Taiwan.

@@

# Methods

>>>>

- start with researches about shape analysis for artefacts and what is the implication.

@@

@@

# Materials and methods

## Archaeological pottery from Kiwulan

Kiwulan is an late Iron Age site situated on a hill near a riverside at the northern margin of Yilan, northeast Taiwan. Yilan is characterized by a triangular alluvial plain facing eastwards the Pacific and mountains on three other sides. Kiwulan contains two archaeological layers, Lower Culture Layer (1300 BP to 800 BP) and Upper Culture Layer (600 BP to 100 BP, i.e. AD 1400-1900), with a sterile deposit in between. The Upper Culture Layer is the main focus in this paper which covers 500 years from the late Iron Age to the historical period. This time period experienced frequent foreign contacts including the European colonial presence in the 17th century and great waves of Chinese immigrant since 19th century. The excavation revealed abundant artifacts in which potsherds are the dominate materials throughout the site. Imported ceramics from mainland China, stonewares, and ornament elements such as beads were usually found in the Upper Culture Layer that indicates the prosperous international trading activities in the 17th century during the European presence in Taiwan. In addition to artifacts, features were also found, such as burials, middens, and postholes with in-situ posts that explains Kiwulan was a settlement site.
